---
title: "Assignment 02, Question 3"
author: "Jeff Nguyen"
date: "February 24, 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, envir=.GlobalEnv)
```

**University of Southern California**  
**Marshall School of Business**  
**FBE 543 Forecasting and Risk Analysis**
  
Student Name: Ngoc Son (Jeff) Nguyen  


# **Question 3**
  
We have the log linear regression equation:
  
\begin{equation}
  \begin{aligned}
    \ln{Q_{it}} =& 915 - 1.5\ln{P_{it}} - .3\ln{P_{ct}} + .9\ln{P_{st}} + .009\ln{Y_t} + .045\ln{AD_t}  \\
  \end{aligned}
\end{equation}
  
```{r}
beta_Pi <- -1.5
beta_Pc <- -.3
beta_Ps <- .9
beta_Y <- .009
beta_AD <- .045
Pi2013 <- 200  # Average market price of (i) in 2013
Qi2013 <- 55000 # Average quantity demanded of (i) in 2013
Y2013 <- 120000 # Average income in 2013
```

## a. With respect to price, what kind of good is $(i)$. For a high demand season, should the company increase the price or decrease it?
  
To determine what kind of good is $(i)$ w.r.t. $P_{it}$, we find the price elasticity of ${i}$ as follow:
    
```{r}
epsilon_Pi <- beta_Pi * Pi2013 / Qi2013  # Price elasticity of (i)
```
  
\begin{equation}
  \begin{aligned}
    \epsilon_{P_{it}} =& \hat{\beta}_{P_{it}} \frac{\overline{P}}{\overline{Q_i}}  \\
                      =& (`r beta_Pi`) \frac{`r Pi2013`}{`r Qi2013`}  \\
                      =& `r epsilon_Pi`  \\
  \end{aligned}
\end{equation}
  
Since $|\epsilon_{P_{it}}| < 1$, $(i)$ is a price inelastic (necessity) good, i.e. price increases leads to total revenue increases. Thus, for a high demand season, the company should increase the price to increase their total revenue.
  
## b. With respect to income, what kind of good is $(i)$.
  
To determine what kind of good is $(i)$ w.r.t. $Y_{t}$, we find the income elasticity of ${i}$ as follow:
    
```{r}
epsilon_Y <- beta_Y * Y2013 / Qi2013  # Income elasticity of (i)
```
  
\begin{equation}
  \begin{aligned}
    \epsilon_{Y_t} =& \hat{\beta}_{Y_t} \frac{\overline{Y}}{\overline{Q_i}}  \\
                      =& (`r beta_Y`) \frac{`r Y2013`}{`r Qi2013`}  \\
                      =& `r epsilon_Y`  \\
  \end{aligned}
\end{equation}
  
Since $\epsilon_{Y_t} < 1$, thus $(i)$ is an income inelastic good (necessity), i.e. a jump in income is less than proportionate than the increase in the quantity demanded.
  
## c. Is the amount of money spent on advertising right amount? Should the company spend more on advertising or less? Why?

```{r}
beta_AD <- .045
AD2013 <- 2800000  # Ad spent for 2013

epsilon_AD <- beta_AD * AD2013 / Qi2013  # Advertising elasticity of demand in 2013
```
  
We have the advertising elasticity of demand as follow:
  
\begin{equation}
  \begin{aligned}
    \epsilon_{AD_t} =& \hat{\beta}_{AD_t} \frac{\overline{AD}}{\overline{Q_i}}  \\
                      =& (`r beta_AD`) \frac{`r AD2013`}{`r Qi2013`}  \\
                      =& `r epsilon_AD`  \\
  \end{aligned}
\end{equation}
  
We can see that quantity demanded of $(i)$ is advertising elastic so the company should spend more as it would lead to revenue increases.
  
The optimum advertising amount in 2013 is:
  
```{r}
AD_2013Optimum <- Pi2013 * Qi2013 * epsilon_AD / abs(epsilon_Pi)
```
  
\begin{equation}
  \begin{aligned}
    AD_{2013}^* =& P_{it} Q_{it} \frac{\epsilon_{AD}}{|\epsilon_{P_{it}}|}  \\
              =& (`r Pi2013`) (`r Qi2013`) \frac{(`r epsilon_AD`)}{|(`r epsilon_Pi`)|}  \\
              =& `r AD_2013Optimum`  \\
  \end{aligned}
\end{equation}
  
## d. What is your forecast of the quantity demanded for 2014?
  
```{r}
# Changes from 2013 to 2014
delta_Pi <- .05  # Price of good (i) increases 5%
delta_Y <- .04  # Income growth 4 %
delta_Ps <- .1  # 10% tax add on to price of substitute good

delta_Qi <- beta_Pi*delta_Pi + beta_Ps*delta_Ps + beta_Y*delta_Y
```
  
We have the forecast $\%$ change in $Q_i$ for 2014 as follow:
  
\begin{equation}
  \begin{aligned}
    \%\Delta Q_i =& \hat{\beta}_{P_{it}} \Delta P_{it} + \hat{\beta}_{P_{st}} \Delta P_{st} + \hat{\beta}_{Y_t} \Delta Y_t  \\
                 =& (`r beta_Pi`)(`r delta_Pi`) + (`r beta_Ps`)(`r delta_Ps`) + (`r beta_Y`)(`r delta_Y`)  \\
                 =& `r delta_Qi`  \\
                 =& `r delta_Qi*100` \%
  \end{aligned}
\end{equation}
  
```{r}
Qi2014 <- Qi2013*(1+delta_Qi)
```
  
Thus, the quantity demanded of $(i)$ increases by $`r delta_Qi*100` \%$ from 2013 to 2014 or $`r Qi2014`$.
  
## e. What is the optimum amount of advertiising spending for 2014?
  
The optimum advertising amount in 2013 is:
  
```{r}
Pi2014 <- Pi2013 * (1+delta_Pi)  # Price of (i) in 2014

AD_2014Optimum <- Pi2014 * Qi2014 * epsilon_AD / abs(epsilon_Pi)
```
  
\begin{equation}
  \begin{aligned}
    AD_{2014}^* =& P_{it} Q_{it} \frac{\epsilon_{AD}}{|\epsilon_{P_{it}}|}  \\
              =& (`r Pi2014`) (`r Qi2014`) \frac{(`r epsilon_AD`)}{|(`r epsilon_Pi`)|}  \\
              =& `r AD_2014Optimum`  \\
  \end{aligned}
\end{equation}