---
title: "Assignment 01, Question 3"
author: "Jeff Nguyen"
date: "February 5, 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, envir=.GlobalEnv)
```

**University of Southern California**  
**Marshall School of Business**  
**FBE 543 Forecasting and Risk Analysis**
  

Student Name: Ngoc Son (Jeff) Nguyen  


# **Question 3**
  
To test whether the pre-split changes in the stock price of TSLA was due to random fluctuations of the market due to the leak of stock split information, we complete an event study statistical analysis. Our assumption is that the stock market is efficient and thus, stock prices adjust quickly to news and information in the market. Steps:
  
1. Download data from Yahoo Finance. The following tickers are downloaded: TSLA, GSPC (representing the S&P 500).
  
2. Define two sets of data for Estimation Window and Event Window. The event happens on August 12, 2020.
  
3. Run a linear regression using Estimation Window.
  
4. Do an ex-post forecast of TSLA daily returns using data for the event period.
  
5. Compute the test statistics: abnormal return (AR) and the t-test statistics on the significance of AR
  
6. Concluding remarks
  
## **Event Study**

### 1. Downloading data:
  
```{r}
library(quantmod)

# Set start date and end date of data
start_date <- "2019-08-01"
end_date <- "2020-10-31"
event_date <- "2020-08-12"

# Get data
getSymbols("TSLA", src = "yahoo", from = start_date, to = end_date)
getSymbols("^GSPC", src = "yahoo", , from = start_date, to = end_date) # S&P 500

# Get adjusted returns data
rTSLA <- diff(log(TSLA$TSLA.Adjusted))
rGSPC <- diff(log(GSPC$GSPC.Adjusted))
```
  
Observe TSLA and S&P 500 for the period:
  
```{r}
# Initialize xts objects contain adjusted price for S&P 500 and AAPL and merge
gspc_xts <- as.xts(GSPC[,"GSPC.Adjusted"])
tsla_xts <- as.xts(TSLA[,"TSLA.Adjusted"])
price_compare <- merge.xts(gspc_xts, tsla_xts)

# Graph monthly AAPL and monthly S&P500 on one coordinate system
# Plot S&P 500
plot(as.zoo(price_compare[, "GSPC.Adjusted"]), screens = 1, 
     main = "S&P 500 and TSLA Adjusted Price Overlay", 
     xlab = "Year", ylab = "Price", col = "Red")

# Keep working on the same plot
par(new = TRUE)

# Plot AAPL and suppress axis value
plot(as.zoo(price_compare[, "TSLA.Adjusted"]), 
     screens = 1, 
     xaxt = "n", yaxt = "n", 
     xlab = "", ylab = "", 
     col = "Blue")

# Add right-handed axis to display AAPL price
axis(4)

# Add legend
legend("topleft", 
       c("S&P 500 Adjusted Price", "TSLA Adjusted Price"), 
       lty = 1:1, 
       cex = 0.75, 
       fill = c("red", "blue"))
```
  
For the period, we can observe TSLA price trend tracks that of the S&P 500 closely with S&P500 experience more volatility at the March 2020 global pandemic shutdown.

### 2. Define two set of data:
  
```{r}
# Data File
data_ret <- data.frame(rTSLA, rGSPC)
colnames(data_ret) <- c('rTSLA', 'rGSPC')

# Define two set of data
data_ret_estimation <- data_ret[4:255,]
cat("The number of trading day in estimation data set is", nrow(data_ret_estimation))

data_ret_test <- data_ret[256:266,]
cat("The number of trading day in test data set is", nrow(data_ret_test))
data_ret_test
```
  
### 3. Run a linear regression using Estimation Window:
  
We attempt to fit an equation of a line: $Y = \alpha_{Jensen} + \beta * X + \epsilon$  

Hypothesis for regression:  
\begin{equation}
  \begin{aligned}
    H_0: \alpha =& 0  \\
    H_a: \alpha \neq& 0  \\
    \text{and}  \\
    H_0: \beta =& 0  \\
    H_a: \beta \neq& 0  \\
  \end{aligned}
\end{equation}
  
```{r}
reg <- lm(rTSLA~rGSPC, data=data_ret_estimation)

summary(reg)
```
  
The estimated equation is $rTSLA = .006730 + 1.181994*rGSPC$, where the $p_{value}$ for the intercept $.0143 < .05$.  
Therefore, we reject the null hypothesis at 95% confidence level that the intercept $\alpha_{Jensen}$ statistically is no different from 0. Thus, we reject the null hypothesis $H_0: \alpha = 0$ and accept the null hypothesis $H_a: \alpha \neq 0$.    
  
Similarly, the $p_{value}$ for $\beta$ is $2e-16 < .05$, implying that the coefficient $\beta$ statistically is significant at 95% or more, and we reject the null hypothesis $H_0: \beta = 0$ and accept the alternative hypothesis $H_a: \beta \neq 0$.  

Goodness of Fit:  
Through inspection of the linear regression result, we observe the $R^2 = .2539$ value to not be close to 1 at all. $R^2 = .2539$ implies that $25.39\%$ of the variations in daily returns of TSLA is explained by that or the S&P 500.
  
Standard Error of Regression:  
We can see that the Standard Error of Regression is $S.E. = .04329$.
  
We have the mean daily returns of TSLA for the period to be $mean(na.omit(rTSLA))$.
  
From this, we can calculate the forecasting efficiency statistic to be: 
\begin{equation}
  \begin{aligned}
    \frac{S.E.}{\overline Y} =& \frac{.04329}{`r mean(na.omit(rTSLA))`}   \\
                             =& `r round(.04329*100/mean(na.omit(rTSLA)),2)`\% > 10\%  \\
  \end{aligned}
\end{equation}
This statistic implies that the linear model is not a good forecasting model.  

Thus, upon exploring the goodness of fit and standard error of regression, we can see that daily returns of TSLA and S&P500 do not follow a linear model.
  
### 4. Do an ex-post forecast of TSLA daily returns using data for the event period.
  
```{r}
library(forecast)

pred <- predict(reg, newdata=data_ret_test, se.fit = TRUE)
pred
```

### 5. Compute the test statistics: abnormal return (AR) and the t-test statistics on the significance of AR
  
Calculate abnormal return statistics:
  
```{r}
#Find Abnormal Return
ar_data <- cbind(data_ret_test$rTSLA,  pred$fit)
ar_data

ar <- data_ret_test$rTSLA - pred$fit
ar

ar_plot <- data.frame(ar)
colnames(ar_plot) <- "AR"
chartSeries(ar_plot, 
            name="Abnormal Returns", 
            type='line', 
            theme=chartTheme("white"))
```
  
Do a t-test for the significance of the abnormal returns:
  
```{r}
T_Data <- cbind(ar, pred$se.fit)
T_test <- data.frame(ar/pred$se.fit)
T_test
```
  
Plot the cummulative abnormal return statistics to examine the impact of the event and the efficiency of the market:
  
```{r}
cum <- data.frame(cumsum(ar))
colnames(cum) <- "CAR"
cum

chartSeries(cum, 
            name="Cumulative Abnormal Returns", 
            type='line', 
            theme=chartTheme("white"))
```
  
### 6. Concluding Remarks
  
The event happens on August 12, 2020. We can observe that the event was a surprise to the market as TSLA sold off prior to the event, from Aug 5 to Aug 11 and increased in price substantially after the announcement. Thus, we conclude that the pre-split changes in the stock price was due to random fluctuations of the market and not because of the leak of the stock split information.
