---
title: "Assignment 04, Question 1&2"
author: "Jeff Nguyen"
date: "March 17, 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, envir=.GlobalEnv)
```

**University of Southern California**  
**Marshall School of Business**  
**FBE 543 Forecasting and Risk Analysis**
  
Student Name: Ngoc Son (Jeff) Nguyen  


# **Question 1**
  
Downloading data:  
```{r}
library(quantmod)

# Set start date and end date of data
start_date <- "2018-01-01"
end_date <- "2021-03-17"

# Get data
getSymbols("AAPL", src = "yahoo", from = start_date, to = end_date)
getSymbols("^GSPC", src = "yahoo", , from = start_date, to = end_date) # S&P 500

# Adjusted Prices
adjAAPL <- AAPL$AAPL.Adjusted
adjGSPC <- GSPC$GSPC.Adjusted
```
  
## a. Graph your AAPL against time (scatter diagram). Comment on the existence of time trend, seasonal trend, cyclical trend, autocorrelation, randomness, structural breaks, and outliers.  
  
```{r}
plot(adjAAPL, main="AAPL", col="Blue")
```
  
**Time Trend:** AAPL displays time trend, as price increases over time.  
**Seasonal Trend:** AAPL displays seasonal trend with up and down spikes in price daily.  
**Cyclical Trend:** AAPL is affected by business cycle of peaks and troughs.  
**Autocorrelation:** AAPL rises for some times when they rise and vice versa.  
**Randomness:** Price of AAPL is unpredictable via inspection.  
**Structural Breaks:** AAPL does not experience any structural break during this time frame as price always recover quickly.  
**Outliers:** AAPL price has several outliers during this period (Jan 2019, March 2020) where price fell more than 30% and recovered.  
  
## b. Graph S&P 500. Comment on the existence of time trend, seasonal trend, cyclical trend, autocorrelation, randomness, structural breaks, and outliers.    
  
```{r}
plot(adjGSPC, main="S&P500", col="Red")
```
  
**Time Trend:** S&P 500 displays time trend, as price increases over time.  
**Seasonal Trend:** S&P 500 displays seasonal trend with up and down spikes in price daily.  
**Cyclical Trend:** S&P 500 is affected by business cycle of peaks and troughs.  
**Autocorrelation:** S&P 500 rises for some times when they rise and vice versa.  
**Randomness:** Price of S&P 500 is unpredictable via inspection.  
**Structural Breaks:** S&P 500 does not experience any structural break during this time frame as price always recover quickly.  
**Outliers:** S&P 500 price has several outliers during this period (Jan 2019, March 2020) where price fell more than 30% and recovered.  

## c. Graph your variable against the market index S&P 500 on x-y axis. Comment on the behavior and the relationship between the two variables.  
  
```{r}
# Initialize xts objects contain adjusted price for S&P 500 and AAPL and merge
gspc_xts <- as.xts(GSPC[,"GSPC.Adjusted"])
aapl_xts <- as.xts(AAPL[,"AAPL.Adjusted"])
price_compare <- merge.xts(gspc_xts, aapl_xts)

# Graph monthly AAPL and monthly S&P500 on one coordinate system
# Plot S&P 500
plot(as.zoo(price_compare[, "GSPC.Adjusted"]), screens = 1, 
     main = "S&P 500 and AAPL Adjusted Price Overlay", 
     xlab = "Year", ylab = "Price", col = "Red")

# Keep working on the same plot
par(new = TRUE)

# Plot AAPL and suppress axis value
plot(as.zoo(price_compare[, "AAPL.Adjusted"]), 
     screens = 1, 
     xaxt = "n", yaxt = "n", 
     xlab = "", ylab = "", 
     col = "Blue")

# Add right-handed axis to display AAPL price
axis(4)

# Add legend
legend("topleft", 
       c("S&P 500 Adjusted Price", "AAPL Adjusted Price"), 
       lty = 1:1, 
       cex = 0.75, 
       fill = c("red", "blue"))
```
  
**Remarks**  
**1. Seasonal Trend:**
Both display similar seasonal trend with up and down spikes in price days after days.  
**2. Cyclical Trend:**  
Both is affected by business cycle with similar peaks and troughs. The S&P 500 has more prominent troughs compared to that of AAPL. Especially during the period of March 2020 with the COVID-19 lockdowns started rolling across the countries.  
**3. Auto-correlation:**  
Both behave similar in this regard where both rises for some times when they rise and vice versa.  
**4. Randomness:**  
Price of both is unpredictable via inspection.  
**5. Time:**  
Both follows similar time trend.  
**6. Structural Break:**  
Both does not experience any structural break during this time frame. During the COVID-19 March sell-off, price recovered quickly in a V-shaped fashion.  
**7. Outliers:**  
SPY looks to have several outliers corresponded to the March 2020 sell-off when the indicies retreated 30% in a short period of a few weeks before quickly recovered.  
  
## d. Repeat a to c for the monthly returns to AAPL and S&P 500.  
  
Adjusting data to monthly:  
```{r}
# Get adjusted returns data
rAAPL <- diff(log(to.monthly(AAPL)$AAPL.Adjusted))
rGSPC <- diff(log(to.monthly(GSPC)$GSPC.Adjusted))
```
  
### d.a. Graph your AAPL monthly returns against time (scatter diagram). Comment on the existence of time trend, seasonal trend, cyclical trend, autocorrelation, randomness, structural breaks, and outliers.  
  
```{r}
plot(rAAPL, main="AAPL Monthly Returns", col="Blue")
```
  
**Time Trend:** We can not observe a clear time trend here. 
**Seasonal Trend:** AAPL monthly returns displays seasonal trend with up and down spikes.  
**Cyclical Trend:** We cannot observe a clear cyclical trend that coincides with business cycles here even though there are peaks and troughs.  
**Autocorrelation:** We cannot observe autocorrelation characteristics.  
**Randomness:** AAPL monthly returns can be observed to be unpredictable.   
**Structural Breaks:** We cannot observe any structural break here.  
**Outliers:** AAPl monthly returns have several clear outliers during this period.  
  
### d.b. Graph S&P 500. Comment on the existence of time trend, seasonal trend, cyclical trend, autocorrelation, randomness, structural breaks, and outliers.    
  
```{r}
plot(rGSPC, main="S&P500", col="Red")
```
  
**Time Trend:** We can not observe a clear time trend here. 
**Seasonal Trend:** S&P 500 monthly returns displays seasonal trend with up and down spikes.  
**Cyclical Trend:** We cannot observe a clear cyclical trend that coincides with business cycles here even though there are peaks and troughs.  
**Autocorrelation:** We cannot observe autocorrelation characteristics.  
**Randomness:** S&P 500 monthly returns can be observed to be unpredictable.   
**Structural Breaks:** We cannot observe any structural break here.  
**Outliers:** S&P 500 monthly returns have several clear outliers during this period.  

### d.c. Graph your variable against the market index S&P 500 on x-y axis. Comment on the behavior and the relationship between the two variables.  
  
```{r}
# Initialize xts objects contain adjusted price for S&P 500 and AAPL and merge
rgspc_xts <- as.xts(rGSPC[,"GSPC.Adjusted"])
raapl_xts <- as.xts(rAAPL[,"AAPL.Adjusted"])
monthlyR_compare <- merge.xts(rgspc_xts, raapl_xts)

# Graph monthly AAPL and monthly S&P500 on one coordinate system
# Plot S&P 500
plot(as.zoo(monthlyR_compare[, "GSPC.Adjusted"]), screens = 1, 
     main = "S&P 500 and AAPL Monthly Returns Overlay", 
     xlab = "Year", ylab = "Monthly Returns", col = "Red")

# Keep working on the same plot
par(new = TRUE)

# Plot AAPL and suppress axis value
plot(as.zoo(monthlyR_compare[, "AAPL.Adjusted"]), 
     screens = 1, 
     xaxt = "n", yaxt = "n", 
     xlab = "", ylab = "", 
     col = "Blue")

# Add right-handed axis to display AAPL price
axis(4)

# Add legend
legend("topleft", 
       c("S&P 500 Monthly Returns", "AAPL Monthly Returns"), 
       lty = 1:1, 
       cex = 0.75, 
       fill = c("red", "blue"))
```
  
**Remarks**  
For the most part, both monthly returns of AAPL and S&P 500 follow each other in lock-step saved for a few periods where AAPL monthly returns become more volatile. S&P 500 experienced larger draw down in returns compared to AAPL in the March 2020 "pandemic sell-off".  
  
## e. Compare the risk and return of AAPL with the risk and return to S&P 500.  

Calculations:  
```{r}
# Calculate statistics
AAPL_meanR <- mean(rAAPL, na.rm=TRUE)
GSPC_meanR <- mean(rGSPC, na.rm=TRUE)

AAPL_riskR <- sqrt(var(rAAPL, na.rm=TRUE))
GSPC_riskR <- sqrt(var(rGSPC, na.rm=TRUE))
```
  
**Comparison:**
  
Underlying | Expected Monthly Returns | Risk
---        | ---                      | ---               
S&P 500    | `r GSPC_meanR`           | `r GSPC_riskR`
AAPL       | `r AAPL_meanR`           | `r AAPL_riskR`
  
We can observe AAPL has higher expected monthly returns but also higher risk than the S&P 500.  

## f. Plot histograms of returns to AAPL and returns to S&P 500. Comment on the distribution of the returns.  

```{r}
hist(rAAPL, 
     main='Monthly Returns for AAPL', 
     xlab='Returns', 
     col='blue',
    )
```
  
We can observe that AAPL has a fat left tail (to the downside).  

```{r}
hist(rGSPC, 
     main='Monthly Returns for S&P 500', 
     xlab='Returns', 
     col='red',
    )
```
  
We can observe that compared to AAPL, S&P 500 resembles, but not necessary is, a normal distribution with fairly even tails.  

## g. Test whether the distributions of returns to AAP and returns to S&P 500 are normal or not.  

We run Shapiro-Wilk normality test on AAPL monthly returns:

```{r}
shapiro.test(as.vector(rAAPL))
```
  
We can see that p-value > .05 implying that the distribution of AAPL monthly returns is not significantly different from a normal distribution, and thus, we can assume normality.  

Similar with S&P 500:  

```{r}
shapiro.test(as.vector(rGSPC))
```
  
Similarly, we can see that the distribution of S&P 500 monthly returns is not significantly different from a normal distribution, and thus, we can assume normality.  

## h. Fit $MA(5)$ and $MA(9)$ on AAPL data and compare the accuracy criterion of the fits.  
  
Fitting $MA(5)$:  

```{r}
library(forecast)

ma5 <- ma(adjAAPL, order=5)  # Get MA5 in ts format
ma9 <- ma(adjAAPL, order=9)  # Get MA9 in ts format
ma5 <- as.xts(ma5)  # Convert to xts
ma9 <- as.xts(ma9)  # Convert to xts
index(ma5) <- index(adjAAPL)  # Replace MA5 index with adjAAPL index
index(ma9) <- index(adjAAPL)  # Replace MA9 index with adjAAPL index

# Merge adjAAPL and its MA5 to a single object
adjAAPL_wMA <- na.omit(merge.xts(adjAAPL, ma5, ma9))

plot(as.zoo(adjAAPL_wMA[, "AAPL.Adjusted"]), screens = 1, 
     main = "AAPL and its 5 day MA Overlay", 
     xlab = "Year", ylab = "Price", col = "blue")

par(new=TRUE)

# Plot MA5 and suppress axis value
plot(as.zoo(adjAAPL_wMA[, "ma5"]), 
     screens = 1,
     lty=2,
     xaxt = "n", yaxt = "n", 
     xlab = "", ylab = "", 
     col = "orange")

par(new=TRUE)

# Plot MA9 and suppress axis value
plot(as.zoo(adjAAPL_wMA[, "ma9"]), 
     screens = 1,
     lty=2,
     xaxt = "n", yaxt = "n", 
     xlab = "", ylab = "", 
     col = "green")

# Add legend
legend("topleft", 
       c("AAPL Daily Price", "AAPL Daily MA5", "AAPL Daily MA9"), 
       lty=c(1,2,2),
       cex = 0.75, 
       fill = c("blue", "orange", "green"))
```
  
We can observe that both MAs smooth out the actual price graph and remove some of its sharp movements.  
  
## i. Fit WMA(5) on AAPL data and compare its accuracy of the fit with MA(5) and MA(9)  
  
Add WMA(5)  
```{r}
library(pracma)

wma5 <- movavg(adjAAPL, n=5, type="w")  # Calculate WMA5
adjAAPL_wWMA <- merge.xts(adjAAPL, wma5)  # Merge WMA5 to AAPL daily price xts object

plot(as.zoo(adjAAPL_wWMA[, "AAPL.Adjusted"]), screens = 1, 
     main = "AAPL and its 5 day WMA Overlay", 
     xlab = "Year", ylab = "Price", col = "blue")

par(new=TRUE)

# Plot MA5 and suppress axis value
plot(as.zoo(adjAAPL_wWMA[, "wma5"]), 
     screens = 1,
     lty=2,
     xaxt = "n", yaxt = "n", 
     xlab = "", ylab = "", 
     col = "orange")

# Add legend
legend("topleft", 
       c("AAPL Daily Price", "AAPL Daily WMA5"), 
       lty=c(1,2),
       cex = 0.75, 
       fill = c("blue", "orange"))
```
  
We can abserve that WMA5 fits daily price a lot better than MA5 and MA9.
  
## j. Fit simple ES to AAPL data and compare its accuracy of the fit with MA(5) and MA(9).  

Simple ES  
```{r}
#Simple Exponential Smoothing with 1 period ahead
fit1 <- ses(adjAAPL, alpha=0.2, initial="simple", h=1)
es <- as.xts(fitted(fit1))
index(es) <- index(adjAAPL)
adjAAPL_wES <- merge.xts(adjAAPL, es)  # Merge ES to AAPL daily price xts object

plot(as.zoo(adjAAPL_wES[, "AAPL.Adjusted"]), screens = 1, 
     main = "AAPL and its Exponential Smoothing Overlay", 
     xlab = "Year", ylab = "Price", col = "blue")

par(new=TRUE)

# Plot ES and suppress axis value
plot(as.zoo(adjAAPL_wES[, "es"]), 
     screens = 1,
     lty=2,
     xaxt = "n", yaxt = "n", 
     xlab = "", ylab = "", 
     col = "chocolate")

# Add legend
legend("topleft", 
       c("AAPL Daily Price", "AAPL Daily ES"), 
       lty=c(1,2),
       cex = 0.75, 
       fill = c("blue", "chocolate"))
```
  
We can observe that ES5 also fit daily price quite well, but poorer than WMA5.
  
## k. Do a one-period ahead forecasting of AAPL price for using simple ES model.  
  
```{r}
fit1
```
  
## l. Fit Holt-Winter ES to AAPL data and compare its accuracy of the fit with MA(5) and MA(9).  


```{r}
#Simple Exponential Smoothing with 1 period ahead
holt1 <- holt(adjAAPL, alpha=0.8, beta=0.2, initial="simple", exponential=TRUE, h=3) 
holt1ES <- as.xts(fitted(holt1))
index(holt1ES) <- index(adjAAPL)

# Merge Holt ES to AAPL daily price xts object
adjAAPL_wHoltES <- merge.xts(adjAAPL, holt1ES)

plot(as.zoo(adjAAPL_wHoltES[, "AAPL.Adjusted"]), screens = 1, 
     main = "AAPL and its Holt-Winters ES Overlay", 
     xlab = "Year", ylab = "Price", col = "blue")

par(new=TRUE)

# Plot ES and suppress axis value
plot(as.zoo(adjAAPL_wHoltES[, "holt1ES"]), 
     screens = 1,
     lty=2,
     xaxt = "n", yaxt = "n", 
     xlab = "", ylab = "", 
     col = "chocolate")

# Add legend
legend("topleft", 
       c("AAPL Daily Price", "AAPL Daily HoltES"), 
       lty=c(1,2),
       cex = 0.75, 
       fill = c("blue", "chocolate"))
```
  
We can oberve that Holt-ES model provides the best fit thus far.  

## m.Do a 3-period ahead forecasting of AAPL price using Holt-Winter ES model.  

```{r}
holt1
```
  
# Question 2  
  
## a. Given that the damping effect, $\alpha = .05$, do next two-periods ES forecast of the prices for Nov and Dec of 2014.  

We have the general exponential smoothing model:  
\begin{equation}
  \begin{aligned}
    Y_{t+1} =& Y_t^F + \alpha (Y_t - Y_t^F)  \\
  \end{aligned}
\end{equation}
  

Plug in:

```{r}
y_Nov14F <- 101.62 + .05*(97.67 - 101.62)
```
\begin{equation}
  \begin{aligned}
    Y_{Nov14}^F =& Y_{t+1}  \\
                =& Y_t^F + \alpha (Y_t - Y_t^F)  \\
                =& Y_{Oct14}^F + \alpha (Y_{Oct14} - Y_{Oct14}^F)  \\
                =& 101.62 + .05*(97.67 - 101.62)  \\
                =& `r y_Nov14F`  \\
  \end{aligned}
\end{equation}
  
\begin{equation}
  \begin{aligned}
    Y_{Dec14}^F =& Y_{t+1}  \\
                =& Y_t^F + \alpha (Y_t - Y_t^F)  \\
                =& Y_{Nov14}^F + \alpha (Y_{Nov14} - Y_{Nov14}^F)  \\
  \end{aligned}
\end{equation}
  
**Remarks** For the period of December 2014, it is not possible to make inference based on existing data as the actual closing price for November 2014 $Y_{Nov14}$ is required, i.e. it is not possible doing 2-period ahead forecasting using Exponential Smoothing.  
  
## b. If SE of the forecast is 1.8, write 95% confidence interval for the forecast of November 2014.  
  
```{r}
epsilon_F <- 1.8
```
95% confidence interval:  
  
\begin{equation}
  \begin{aligned}
    Y_{Nov14}^F - 1.96.\epsilon_F \leqslant & Y_{Nov14} \leqslant Y_{Nov14}^F - 1.96.\epsilon_F  \\
    `r y_Nov14F` - 1.96.\epsilon_F \leqslant & Y_{Nov14} \leqslant `r y_Nov14F`` - 1.96.\epsilon_F  \\
    `r y_Nov14F - 1.96*epsilon_F` \leqslant & Y_{Nov14} \leqslant `r y_Nov14F + 1.96*epsilon_F`  \\
  \end{aligned}
\end{equation}