---
title: "Assignment 04, Question 1&2"
author: "Jeff Nguyen"
date: "March 17, 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, envir=.GlobalEnv)
```

**University of Southern California**  
**Marshall School of Business**  
**FBE 543 Forecasting and Risk Analysis**
  
Student Name: Ngoc Son (Jeff) Nguyen  


# **Question 1**
  
Downloading data:  
```{r}
library(quantmod)

# Set start date and end date of data
start_date <- "2018-01-01"
end_date <- "2021-03-17"

# Get data
getSymbols("AAPL", src = "yahoo", from = start_date, to = end_date)
getSymbols("^GSPC", src = "yahoo", , from = start_date, to = end_date) # S&P 500

# Adjusted Prices
adjAAPL <- AAPL$AAPL.Adjusted
adjGSPC <- GSPC$GSPC.Adjusted
```
  
## a. Graph your AAPL against time (scatter diagram). Comment on the existence of time trend, seasonal trend, cyclical trend, autocorrelation, randomness, structural breaks, and outliers.  
  
```{r}
plot(adjAAPL, main="AAPL", col="Blue")
```
  
**Time Trend:** AAPL displays time trend, as price increases over time.  
**Seasonal Trend:** AAPL displays seasonal trend with up and down spikes in price daily.  
**Cyclical Trend:** AAPL is affected by business cycle of peaks and troughs.  
**Autocorrelation:** AAPL rises for some times and they rise and vice versa.  
**Randomness:** Price of AAPL is unpredictable via inspection.  
**Structural Breaks:** AAPL does not experience any structural break during this time frame as price always recover quickly.  
**Outliers:** AAPL price has several outliers during this period (Jan 2019, March 2020) where price fell more than 30% and recovered.  
  
## b. Graph S&P 500. Comment on the existence of time trend, seasonal trend, cyclical trend, autocorrelation, randomness, structural breaks, and outliers.    
  
```{r}
plot(adjGSPC, main="S&P500", col="Red")
```
  
**Time Trend:** S&P 500 displays time trend, as price increases over time.  
**Seasonal Trend:** S&P 500 displays seasonal trend with up and down spikes in price daily.  
**Cyclical Trend:** S&P 500 is affected by business cycle of peaks and troughs.  
**Autocorrelation:** S&P 500 rises for some times and they rise and vice versa.  
**Randomness:** Price of S&P 500 is unpredictable via inspection.  
**Structural Breaks:** S&P 500 does not experience any structural break during this time frame as price always recover quickly.  
**Outliers:** S&P 500 price has several outliers during this period (Jan 2019, March 2020) where price fell more than 30% and recovered.  

## c. Graph your variable against the market index S&P 500 on x-y axis. Comment on the behavior and the relationship between the two variables.  
  
```{r}
# Initialize xts objects contain adjusted price for S&P 500 and AAPL and merge
gspc_xts <- as.xts(GSPC[,"GSPC.Adjusted"])
aapl_xts <- as.xts(AAPL[,"AAPL.Adjusted"])
price_compare <- merge.xts(gspc_xts, aapl_xts)

# Graph monthly AAPL and monthly S&P500 on one coordinate system
# Plot S&P 500
plot(as.zoo(price_compare[, "GSPC.Adjusted"]), screens = 1, 
     main = "S&P 500 and AAPL Adjusted Price Overlay", 
     xlab = "Year", ylab = "Price", col = "Red")

# Keep working on the same plot
par(new = TRUE)

# Plot AAPL and suppress axis value
plot(as.zoo(price_compare[, "AAPL.Adjusted"]), 
     screens = 1, 
     xaxt = "n", yaxt = "n", 
     xlab = "", ylab = "", 
     col = "Blue")

# Add right-handed axis to display AAPL price
axis(4)

# Add legend
legend("topleft", 
       c("S&P 500 Adjusted Price", "AAPL Adjusted Price"), 
       lty = 1:1, 
       cex = 0.75, 
       fill = c("red", "blue"))
```
  
**Remarks**  
**1. Seasonal Trend:**
Both display similar seasonal trend with up and down spikes in price days after days.  
**2. Cyclical Trend:**  
Both is affected by business cycle with similar peaks and troughs. The S&P 500 has more prominent troughs compared to that of AAPL. Especially during the period of March 2020 with the COVID-19 lockdowns started rolling across the countries.  
**3. Auto-correlation:**  
Both behave similar in this regard where both rises for some times when they rise and vice versa.  
**4. Randomness:**  
Price of both is unpredictable via inspection.  
**5. Time:**  
Both follows similar time trend.  
**6. Structural Break:**  
Both does not experience any structural break during this time frame. During the COVID-19 March sell-off, price recovered quickly in a V-shaped fashion.  
**7. Outliers:**  
SPY looks to have several outliers corresponded to the March 2020 sell-off when the indicies retreated 30% in a short period of a few weeks before quickly recovered.  
  
## d. Repeat a to c for the monthly returns to AAPL and S&P 500.  
  
Adjusting data to monthly:  
```{r}
# Get adjusted returns data
rAAPL <- diff(log(to.monthly(AAPL)$AAPL.Adjusted))
rGSPC <- diff(log(to.monthly(GSPC)$GSPC.Adjusted))
```
  
### d.a. Graph your AAPL monthly returns against time (scatter diagram). Comment on the existence of time trend, seasonal trend, cyclical trend, autocorrelation, randomness, structural breaks, and outliers.  
  
```{r}
plot(rAAPL, main="AAPL Monthly Returns", col="Blue")
```
  
**Time Trend:** We can not observe a clear time trend here. 
**Seasonal Trend:** AAPL monthly returns displays seasonal trend with up and down spikes.  
**Cyclical Trend:** We cannot observe a clear cyclical trend that coincides with business cycles here even though there are peaks and troughs.  
**Autocorrelation:** We cannot observe autocorrelation characteristics.  
**Randomness:** AAPL monthly returns can be observed to be unpredictable.   
**Structural Breaks:** We cannot observe any structural break here.  
**Outliers:** AAPl monthly returns have several clear outliers during this period.  
  
### d.b. Graph S&P 500. Comment on the existence of time trend, seasonal trend, cyclical trend, autocorrelation, randomness, structural breaks, and outliers.    
  
```{r}
plot(rGSPC, main="S&P500", col="Red")
```
  
**Time Trend:** We can not observe a clear time trend here. 
**Seasonal Trend:** S&P 500 monthly returns displays seasonal trend with up and down spikes.  
**Cyclical Trend:** We cannot observe a clear cyclical trend that coincides with business cycles here even though there are peaks and troughs.  
**Autocorrelation:** We cannot observe autocorrelation characteristics.  
**Randomness:** S&P 500 monthly returns can be observed to be unpredictable.   
**Structural Breaks:** We cannot observe any structural break here.  
**Outliers:** S&P 500 monthly returns have several clear outliers during this period.  

### d.c. Graph your variable against the market index S&P 500 on x-y axis. Comment on the behavior and the relationship between the two variables.  
  
```{r}
# Initialize xts objects contain adjusted price for S&P 500 and AAPL and merge
rgspc_xts <- as.xts(rGSPC[,"GSPC.Adjusted"])
raapl_xts <- as.xts(rAAPL[,"AAPL.Adjusted"])
monthlyR_compare <- merge.xts(rgspc_xts, raapl_xts)

# Graph monthly AAPL and monthly S&P500 on one coordinate system
# Plot S&P 500
plot(as.zoo(monthlyR_compare[, "GSPC.Adjusted"]), screens = 1, 
     main = "S&P 500 and AAPL Monthly Returns Overlay", 
     xlab = "Year", ylab = "Monthly Returns", col = "Red")

# Keep working on the same plot
par(new = TRUE)

# Plot AAPL and suppress axis value
plot(as.zoo(monthlyR_compare[, "AAPL.Adjusted"]), 
     screens = 1, 
     xaxt = "n", yaxt = "n", 
     xlab = "", ylab = "", 
     col = "Blue")

# Add right-handed axis to display AAPL price
axis(4)

# Add legend
legend("topleft", 
       c("S&P 500 Monthly Returns", "AAPL Monthly Returns"), 
       lty = 1:1, 
       cex = 0.75, 
       fill = c("red", "blue"))
```
  
**Remarks**  
For the most part, both monthly returns of AAPL and S&P 500 follow each other in lock-step saved for a few periods where AAPL monthly returns become more volatile. S&P 500 experienced larger draw down in returns compared to AAPL in the March 2020 "pandemic sell-off".  
  
## e. Compare the risk and return of AAPL with the risk and return to S&P 500.  

Calculations:  
```{r}
# Calculate statistics
AAPL_meanR <- mean(rAAPL, na.rm=TRUE)
GSPC_meanR <- mean(rGSPC, na.rm=TRUE)

AAPL_riskR <- sqrt(var(rAAPL, na.rm=TRUE))
GSPC_riskR <- sqrt(var(rGSPC, na.rm=TRUE))
```
  
**Comparison:**
  
Underlying | Expected Monthly Returns | Risk
---        | ---                      | ---               
S&P 500    | `r GSPC_meanR`           | `r GSPC_riskR`
AAPL       | `r AAPL_meanR`           | `r AAPL_riskR`
  
We can observe AAPL has higher expected monthly returns but also higher risk than the S&P 500.  

## f. Plot histograms of returns to AAPL and returns to S&P 500. Comment on the distribution of the returns.  

```{r}
hist(rAAPL, 
     main='Monthly Returns for AAPL', 
     xlab='Returns', 
     col='blue',
    )
```
  
We can observe that AAPL has a fat left tail (to the downside).  

```{r}
hist(rGSPC, 
     main='Monthly Returns for S&P 500', 
     xlab='Returns', 
     col='red',
    )
```
  
We can observe that compared to AAPL, S&P 500 resembles, but not necessary is, a normal distribution with fairly even tails.  

## g. Test whether the distributions of returns to AAP and returns to S&P 500 are normal or not.  