---
title: "Assignment 04, Question 3&4"
author: "Jeff Nguyen"
date: "March 26, 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, envir=.GlobalEnv)
```

**University of Southern California**  
**Marshall School of Business**  
**FBE 543 Forecasting and Risk Analysis**
  
Student Name: Ngoc Son (Jeff) Nguyen  


# **Question 3**
  
## a.  
  
\begin{equation}
  \begin{aligned}
    Y_t =& 2.5 + .35Y_{t-1} + \epsilon_t  \\
  \end{aligned}
\end{equation}

- **ARIMA model:** $ARIMA(1,0,0)$.  
- **Stationarity:** Stationary as $\sum_{i=1}^n |\alpha_i| = .35 < 1$.  
- **Invertibility:** N/A.  

## b.  
  
\begin{equation}
  \begin{aligned}
    Y_t =& 4.5 - 1.5Y_{t-1} + \epsilon_t - .5\epsilon_{t-1}  \\
  \end{aligned}
\end{equation}
  
- **ARIMA model:** $ARIMA(1,0,1)$.  
- **Stationarity:** Non-Stationary as $\sum_{i=1}^n |\alpha_i| = |1.5| > 1$.  
- **Invertibility:** Invertible as $\sum_{i=1}^n \beta_i| = .5 < 1$.  
  
## c.  
  
\begin{equation}
  \begin{aligned}
    Y_t =& 1.2 - .75Y_{t-1} + .3Y_{t-2} + \epsilon_t  \\
  \end{aligned}
\end{equation}
  
- **ARIMA model:** $ARIMA(2,0,0)$.  
- **Stationarity:** Non-Stationary as $\sum_{i=1}^n |\alpha_i| = |-.75| + |.3| = `r .75 + .3` > 1$.  
- **Invertibility:** N/A.  
  
  
## d.  
  
\begin{equation}
  \begin{aligned}
    Y_t =& 2.5 - .95Y_{t-1} + \epsilon_t - .5\epsilon_{t-1} - .2\epsilon{t-2}  \\
  \end{aligned}
\end{equation}
  
- **ARIMA model:** $ARIMA(1,0,2)$.  
- **Stationarity:** Stationary as $\sum_{i=1}^n |\alpha_i| = `r .95` < 1$.  
- **Invertibility:** Invertible as $\sum_{i=1}^n |\beta_i| = |-.5| + |-.2| = `r .5 + .2` < 1$.   
  
  
## e.  
  
\begin{equation}
  \begin{aligned}
    Y_t =& .52 - 1.2Y_{t-1} + \epsilon_t + .2\epsilon_{t-1}  \\
  \end{aligned}
\end{equation}
  
- **ARIMA model:** $ARIMA(1,0,1)$.  
- **Stationarity:** Non-Stationary as $\sum_{i=1}^n |\alpha_i| = `r 1.2` > 1$.  
- **Invertibility:** Invertible as $\sum_{i=1}^n |\beta_i| = |.2| = `r .2` < 1$.   
  
## f.  
  
\begin{equation}
  \begin{aligned}
    DY_t =& 1.2DY_{t-1} + \epsilon_t  \\
  \end{aligned}
\end{equation}
  
- **ARIMA model:** $ARIMA(1,1,0)$.  
- **Stationarity:** Non-Stationary as $\sum_{i=1}^n |\alpha_i| = `r 1.2` > 1$.  
- **Invertibility:** N/A.   
  
  
## g.  
  
\begin{equation}
  \begin{aligned}
    DY_t =& .42DY_{t-1} + \epsilon_t - .6\epsilon_{t-1}  \\
  \end{aligned}
\end{equation}
  
- **ARIMA model:** $ARIMA(1,1,1)$.  
- **Stationarity:** Stationary as $\sum_{i=1}^n |\alpha_i| = `r .42` < 1$.  
- **Invertibility:** Invertible as $\sum_{i=1}^n |\beta_i| = |-.6| = `r .6` < 1$.   
  
  
## h.  
  
\begin{equation}
  \begin{aligned}
    DY_t =& .62Y_{t-1} + \epsilon_t - .6\epsilon_{t-1}  \\
  \end{aligned}
\end{equation}
  
- **ARIMA model:** Incorrect model as $LHS$ and $RHS$ does not have similar order difference term.  
- **Stationarity:** N/A.  
- **Invertibility:** N/A.   
  
  
## i.  
  
\begin{equation}
  \begin{aligned}
    Y_t =& 2.5 + .95DY_{t-1} + \epsilon_t  \\
  \end{aligned}
\end{equation}
  
- **ARIMA model:** Incorrect model as $LHS$ and $RHS$ does not have similar order difference term.  
- **Stationarity:** N/A.  
- **Invertibility:** N/A.   
  
  
## j.  
  
\begin{equation}
  \begin{aligned}
    Y_t =& 1.6 + \epsilon_t - .6\epsilon_{t-1}  \\
  \end{aligned}
\end{equation}
  
- **ARIMA model:** $ARIMA(0,0,1)$.  
- **Stationarity:** Stationary as $\sum_{i=1}^n |\beta_i| = |-.6| = `r .6` < 1$.  
- **Invertibility:** Invertible as $\sum_{i=1}^n |\beta_i| = |-.6| = `r .6` < 1$.   
  
# Question 4  
  
## a. Do a three-period ahead forecasting using the given initial values and statistics. Write 95% confidence interval for each forecast.  

### a)  

\begin{equation}
  \begin{aligned}
    Y_t =& 1.6 + .75Y_{t-1} + \epsilon_t  \\
    \text{given:}&  \\
    y_0 =& 2 \\
    \sigma^2 =& 1.21 \\
  \end{aligned}
\end{equation}
  
```{r}
y_0 <- 2
sigmaSq <- 1.21

f <- function(y_t) {1.6 + .75*y_t}
y_1 <- f(y_0)
y_2 <- f(y_1)
y_3 <- f(y_2)
```
  
| Forecast Variable      | Forecasted Value | 95% Confidence Interval Equation                               | 95% Confidence Interval                                                                     |
|------------------------|------------------|----------------------------------------------------------------|---------------------------------------------------------------------------------------------|
| $Y_1^F = 1.6 + .75Y_0$ | $`r y_1`$        | $Y_{t+1}^F \pm 1.96\sqrt{\sigma^2}$                            | [`r y_1 - 1.96*sqrt(sigmaSq)`, `r y_1 + 1.96*sqrt(sigmaSq)`]                                |
| $Y_2^F = 1.6 + .75Y_1$ | $`r y_2`$        | $Y_{t+2}^F \pm 1.96\sqrt{\sigma^2(1+\alpha^2)}$                | [`r y_2 - 1.96*sqrt(sigmaSq*(1+.75^2))`,`r y_2 + 1.96*sqrt(sigmaSq*(1+.75^2))`]             |
| $Y_3^F = 1.6 + .75Y_2$ | $`r y_3`$        | $Y_{t+3}^F \pm 1.96\sqrt{\sigma^2(1+\alpha_1^2 + \alpha_1^4)}$ | [`r y_3 - 1.96*sqrt(sigmaSq*(1+.75^2+.75^4))`,`r y_3 + 1.96*sqrt(sigmaSq*(1+.75^2+.75^4))`] |
  

### b)  

\begin{equation}
  \begin{aligned}
    Y_t =& 2.5 + .3Y_{t-1} + \epsilon_t  \\
    \text{given:}&  \\
    y_0 =& 10 \\
    \sigma^2 =& 6.25 \\
  \end{aligned}
\end{equation}
  
```{r}
y_0 <- 10
sigmaSq <- 6.25

f <- function(y_t) {2.5 + .3*y_t}
y_1 <- f(y_0)
y_2 <- f(y_1)
y_3 <- f(y_2)
```
  
| Forecast Variable      | Forecasted Value | 95% Confidence Interval Equation                               | 95% Confidence Interval                                                                     |
|------------------------|------------------|----------------------------------------------------------------|---------------------------------------------------------------------------------------------|
| $Y_1^F = 2.5 + .3Y_0$ | $`r y_1`$        | $Y_{t+1}^F \pm 1.96\sqrt{\sigma^2}$                            | [`r y_1 - 1.96*sqrt(sigmaSq)`, `r y_1 + 1.96*sqrt(sigmaSq)`]                                |
| $Y_2^F = 2.5 + .3Y_1$ | $`r y_2`$        | $Y_{t+2}^F \pm 1.96\sqrt{\sigma^2(1+\alpha^2)}$                | [`r y_2 - 1.96*sqrt(sigmaSq*(1+.75^2))`,`r y_2 + 1.96*sqrt(sigmaSq*(1+.75^2))`]             |
| $Y_3^F = 2.5 + .3Y_2$ | $`r y_3`$        | $Y_{t+3}^F \pm 1.96\sqrt{\sigma^2(1+\alpha_1^2 + \alpha_1^4)}$ | [`r y_3 - 1.96*sqrt(sigmaSq*(1+.75^2+.75^4))`,`r y_3 + 1.96*sqrt(sigmaSq*(1+.75^2+.75^4))`] |
  
  

### c)  

\begin{equation}
  \begin{aligned}
    Y_t =& 1.2 - .2Y_{t-1} + \epsilon_t  \\
    \text{given:}&  \\
    y_0 =& 1.5  \\
    \sigma^2 =& .49  \\
  \end{aligned}
\end{equation}
  
```{r}
y_0 <- 1.5
sigmaSq <- .49

f <- function(y_t) {1.2 - .2*y_t}
y_1 <- f(y_0)
y_2 <- f(y_1)
y_3 <- f(y_2)
```
  
| Forecast Variable      | Forecasted Value | 95% Confidence Interval Equation                               | 95% Confidence Interval                                                                     |
|------------------------|------------------|----------------------------------------------------------------|---------------------------------------------------------------------------------------------|
| $Y_1^F = 1.2 - .2Y_0$ | $`r y_1`$        | $Y_{t+1}^F \pm 1.96\sqrt{\sigma^2}$                            | [`r y_1 - 1.96*sqrt(sigmaSq)`, `r y_1 + 1.96*sqrt(sigmaSq)`]                                |
| $Y_2^F = 1.2 - .2Y_1$ | $`r y_2`$        | $Y_{t+2}^F \pm 1.96\sqrt{\sigma^2(1+\alpha^2)}$                | [`r y_2 - 1.96*sqrt(sigmaSq*(1+.75^2))`,`r y_2 + 1.96*sqrt(sigmaSq*(1+.75^2))`]             |
| $Y_3^F = 1.2 - .2Y_2$ | $`r y_3`$        | $Y_{t+3}^F \pm 1.96\sqrt{\sigma^2(1+\alpha_1^2 + \alpha_1^4)}$ | [`r y_3 - 1.96*sqrt(sigmaSq*(1+.75^2+.75^4))`,`r y_3 + 1.96*sqrt(sigmaSq*(1+.75^2+.75^4))`] |
  
  

### d)  

\begin{equation}
  \begin{aligned}
    Y_t =& 2.5 - .8Y_{t-1} + \epsilon_t  \\
    \text{given:}&  \\
    y_0 =& 6  \\
    \sigma^2 =& 3.69  \\
  \end{aligned}
\end{equation}
  
```{r}
y_0 <- 6
sigmaSq <- 3.69

f <- function(y_t) {2.5 - .8*y_t}
y_1 <- f(y_0)
y_2 <- f(y_1)
y_3 <- f(y_2)
```
  
| Forecast Variable      | Forecasted Value | 95% Confidence Interval Equation                               | 95% Confidence Interval                                                                     |
|------------------------|------------------|----------------------------------------------------------------|---------------------------------------------------------------------------------------------|
| $Y_1^F = 2.5 - .8Y_0$ | $`r y_1`$        | $Y_{t+1}^F \pm 1.96\sqrt{\sigma^2}$                            | [`r y_1 - 1.96*sqrt(sigmaSq)`, `r y_1 + 1.96*sqrt(sigmaSq)`]                                |
| $Y_2^F = 2.5 - .8Y_1$ | $`r y_2`$        | $Y_{t+2}^F \pm 1.96\sqrt{\sigma^2(1+\alpha^2)}$                | [`r y_2 - 1.96*sqrt(sigmaSq*(1+.75^2))`,`r y_2 + 1.96*sqrt(sigmaSq*(1+.75^2))`]             |
| $Y_3^F = 2.5 - .8Y_2$ | $`r y_3`$        | $Y_{t+3}^F \pm 1.96\sqrt{\sigma^2(1+\alpha_1^2 + \alpha_1^4)}$ | [`r y_3 - 1.96*sqrt(sigmaSq*(1+.75^2+.75^4))`,`r y_3 + 1.96*sqrt(sigmaSq*(1+.75^2+.75^4))`] |
  
  

### e)  

\begin{equation}
  \begin{aligned}
    Y_t =& - .5Y_{t-1} + \epsilon_t  \\
    \text{given:}&  \\
    y_0 =& -1.6  \\
    \sigma^2 =& 1.44  \\
  \end{aligned}
\end{equation}
  
```{r}
y_0 <- -1.6
sigmaSq <- 1.44

f <- function(y_t) {-.5*y_t}
y_1 <- f(y_0)
y_2 <- f(y_1)
y_3 <- f(y_2)
```
  
| Forecast Variable      | Forecasted Value | 95% Confidence Interval Equation                               | 95% Confidence Interval                                                                     |
|------------------------|------------------|----------------------------------------------------------------|---------------------------------------------------------------------------------------------|
| $Y_1^F = -.5Y_0$ | $`r y_1`$        | $Y_{t+1}^F \pm 1.96\sqrt{\sigma^2}$                            | [`r y_1 - 1.96*sqrt(sigmaSq)`, `r y_1 + 1.96*sqrt(sigmaSq)`]                                |
| $Y_2^F = -.5Y_1$ | $`r y_2`$        | $Y_{t+2}^F \pm 1.96\sqrt{\sigma^2(1+\alpha^2)}$                | [`r y_2 - 1.96*sqrt(sigmaSq*(1+.75^2))`,`r y_2 + 1.96*sqrt(sigmaSq*(1+.75^2))`]             |
| $Y_3^F = -.5Y_2$ | $`r y_3`$        | $Y_{t+3}^F \pm 1.96\sqrt{\sigma^2(1+\alpha_1^2 + \alpha_1^4)}$ | [`r y_3 - 1.96*sqrt(sigmaSq*(1+.75^2+.75^4))`,`r y_3 + 1.96*sqrt(sigmaSq*(1+.75^2+.75^4))`] |
  
  
## Do a long-run (unconditional) forecasting and write 95% confidence interval.  
  
### a)  
  
\begin{equation}
  \begin{aligned}
    Y_t =& 1.6 + .75Y_{t-1} + \epsilon_t  \\
    \text{given:}&  \\
    y_0 =& 2 \\
    \sigma^2 =& 1.21 \\
  \end{aligned}
\end{equation}
  
```{r}
alpha_0 <- 1.6
alpha_1 <- .75
y_0 <- 2
sigmaSq <- 1.21

yLR <- alpha_0 / (1 - alpha_1)
sigmaSq_yLR <- sigmaSq / (1 - alpha_1^2)
y_1 <- f(y_0)
y_2 <- f(y_1)
y_3 <- f(y_2)
```
  
| Forecast Long Run                          | Forecasted Value | 95% Confidence Interval Equation                      | 95% Confidence Interval                                              |
|--------------------------------------------|------------------|-------------------------------------------------------|----------------------------------------------------------------------|
| $Y_{LR} = \frac{`r alpha_0`}{`r alpha_1`}$ | $`r yLR`$        | $Y_{LR} \pm 1.96\sqrt{\frac{\sigma^2}{1-\alpha_1^2}}$ | [`r yLR - 1.96*sqrt(sigmaSq_yLR)`, `r yLR + 1.96*sqrt(sigmaSq_yLR)`] |
  
### b)  
  
\begin{equation}
  \begin{aligned}
    Y_t =& 2.5 + .3Y_{t-1} + \epsilon_t  \\
    \text{given:}&  \\
    y_0 =& 10 \\
    \sigma^2 =& 6.25 \\
  \end{aligned}
\end{equation}
  
```{r}
alpha_0 <- 2.5
alpha_1 <- .3
y_0 <- 10
sigmaSq <- 6.25

yLR <- alpha_0 / (1 - alpha_1)
sigmaSq_yLR <- sigmaSq / (1 - alpha_1^2)
y_1 <- f(y_0)
y_2 <- f(y_1)
y_3 <- f(y_2)
```
  
| Forecast Long Run                          | Forecasted Value | 95% Confidence Interval Equation                      | 95% Confidence Interval                                              |
|--------------------------------------------|------------------|-------------------------------------------------------|----------------------------------------------------------------------|
| $Y_{LR} = \frac{`r alpha_0`}{`r alpha_1`}$ | $`r yLR`$        | $Y_{LR} \pm 1.96\sqrt{\frac{\sigma^2}{1-\alpha_1^2}}$ | [`r yLR - 1.96*sqrt(sigmaSq_yLR)`, `r yLR + 1.96*sqrt(sigmaSq_yLR)`] |
  
### c)  
  
\begin{equation}
  \begin{aligned}
    Y_t =& 1.2 - .2Y_{t-1} + \epsilon_t  \\
    \text{given:}&  \\
    y_0 =& 1.5  \\
    \sigma^2 =& .49  \\
  \end{aligned}
\end{equation}
  
```{r}
alpha_0 <- 1.2
alpha_1 <- -.2
y_0 <- 1.5
sigmaSq <- .49

yLR <- alpha_0 / (1 - alpha_1)
sigmaSq_yLR <- sigmaSq / (1 - alpha_1^2)
y_1 <- f(y_0)
y_2 <- f(y_1)
y_3 <- f(y_2)
```
  
| Forecast Long Run                          | Forecasted Value | 95% Confidence Interval Equation                      | 95% Confidence Interval                                              |
|--------------------------------------------|------------------|-------------------------------------------------------|----------------------------------------------------------------------|
| $Y_{LR} = \frac{`r alpha_0`}{`r alpha_1`}$ | $`r yLR`$        | $Y_{LR} \pm 1.96\sqrt{\frac{\sigma^2}{1-\alpha_1^2}}$ | [`r yLR - 1.96*sqrt(sigmaSq_yLR)`, `r yLR + 1.96*sqrt(sigmaSq_yLR)`] |
  
### d)  
  
\begin{equation}
  \begin{aligned}
    Y_t =& 2.5 - .8Y_{t-1} + \epsilon_t  \\
    \text{given:}&  \\
    y_0 =& 6  \\
    \sigma^2 =& 3.69  \\
  \end{aligned}
\end{equation}
  
```{r}
alpha_0 <- 2.5
alpha_1 <- -.8
y_0 <- 6
sigmaSq <- 3.69

yLR <- alpha_0 / (1 - alpha_1)
sigmaSq_yLR <- sigmaSq / (1 - alpha_1^2)
y_1 <- f(y_0)
y_2 <- f(y_1)
y_3 <- f(y_2)
```
  
| Forecast Long Run                          | Forecasted Value | 95% Confidence Interval Equation                      | 95% Confidence Interval                                              |
|--------------------------------------------|------------------|-------------------------------------------------------|----------------------------------------------------------------------|
| $Y_{LR} = \frac{`r alpha_0`}{`r alpha_1`}$ | $`r yLR`$        | $Y_{LR} \pm 1.96\sqrt{\frac{\sigma^2}{1-\alpha_1^2}}$ | [`r yLR - 1.96*sqrt(sigmaSq_yLR)`, `r yLR + 1.96*sqrt(sigmaSq_yLR)`] |
  
### e)  
  
\begin{equation}
  \begin{aligned}
    Y_t =& - .5Y_{t-1} + \epsilon_t  \\
    \text{given:}&  \\
    y_0 =& -1.6  \\
    \sigma^2 =& 1.44  \\
  \end{aligned}
\end{equation}
  
```{r}
alpha_0 <- 0
alpha_1 <- -.5
y_0 <- -1.6
sigmaSq <- 1.44

yLR <- alpha_0 / (1 - alpha_1)
sigmaSq_yLR <- sigmaSq / (1 - alpha_1^2)
y_1 <- f(y_0)
y_2 <- f(y_1)
y_3 <- f(y_2)
```
  
| Forecast Long Run                          | Forecasted Value | 95% Confidence Interval Equation                      | 95% Confidence Interval                                              |
|--------------------------------------------|------------------|-------------------------------------------------------|----------------------------------------------------------------------|
| $Y_{LR} = \frac{`r alpha_0`}{`r alpha_1`}$ | $`r yLR`$        | $Y_{LR} \pm 1.96\sqrt{\frac{\sigma^2}{1-\alpha_1^2}}$ | [`r yLR - 1.96*sqrt(sigmaSq_yLR)`, `r yLR + 1.96*sqrt(sigmaSq_yLR)`] |
  
